Host richness increases the occurrence but not the severity of bark beetle-induced tree mortality

### 2021-09-13 Code/Manuscript.Rmd
updating

### Import libraries
ipak(c('tidyverse', 'tidymodels', 'reshape2', 'here', 'rgdal', 'sp', 'raster', 'stars', 'sf', 'prism', 'tigris', "tmap", 'fasterize', 'ggplot2','RColorBrewer','ggsci', 'patchwork', 'colorspace', 'grid', 'knitr',  'kableExtra', 'bookdown',  'pdp', 'ranger', 'VSURF', 'themis', 'foreach'))

@@

Construct random forest models using synthetic minority oversampling technique (SMOTE)

#### MPB

```{r fit MPB RF model, eval=F}
response <- "agentMPB"
vars <- c("BA.PICO", "QMD.PICO", "propBA.PICO")
datBB.c <- df[df$focalhost=="Any",]

splitit <- initial_split(datBB.c, prop=.7, strata=response)
datz.train <- training(splitit)
datz.test <- testing(splitit)
datz.train.cv <- vfold_cv(datz.train, strata=response, v=10)

# set formula
BBoccurrence_recipe <- recipe(formula(paste(response, "~", paste(vars, collapse = " + "))), datz.train)

# set sample weights
wts <- 1/(table(datBB.c[,response])/ max(table(datBB.c[,response])))
wts <- wts/sum(wts)

# set model
rf_model <- rand_forest() %>% set_args( trees=tune(), min_n=tune()) %>% set_engine("ranger", importance = "permutation", probability=T, seed=913, sample.fraction=wts) %>% set_mode("classification")

@@

# tune model
tgrid <- expand.grid(trees = c(100,500), min_n=c(10, 50, 100))
rf_tune_results <- rf_workflow %>% tune_grid(resamples = datz.train.cv,  grid = tgrid, metrics = metric_set(roc_auc, accuracy, sens, spec))

@@

# Use backward selection to identify important variables
#dfo.vsurf <- VSURF::VSURF(datz.train[,vars], factor(datz.train[,response]), RFimplem="ranger", parallel=T, ntree=param_best$trees, min.node.size=param_best$min_n)
#registerDoSEQ()
#vars.final <- unique(c(vars[dfo.vsurf$varselect.pred]))
vars.final <- vars
BBoccurrence_recipe_final <- recipe(formula(paste(response, "~", paste(vars.final, collapse = " + "))), data=datBB.c)

rf_model_final <- rand_forest() %>% set_args(trees=param_best$trees, min_n=param_best$min_n, splitrule="gini") %>% set_engine("ranger", importance = "permutation", probability=T, sample.fraction=wts) %>% set_mode("classification")

rf_workflow_final <- workflow() %>% add_recipe(BBoccurrence_recipe_final) %>% add_model(rf_model_final)
rf_workflow_final_fit<- rf_workflow_final %>% last_fit(splitit, metrics=metric_set(roc_auc, accuracy, sens, spec)) #emulates the process where, after determining the best model, the final fit on the entire training set is needed and is then evaluated on the test set. 
test_performance <- rf_workflow_final_fit %>% collect_metrics()
write.csv(test_performance, here(paste0("Results/RF-testperformance-", response, "-occurrence-2.csv")), row.names=F)

# generate a confusion matrix for test set
test_predictions <- rf_workflow_final_fit %>% collect_predictions()
confusion <- test_predictions %>% conf_mat(response,estimate = .pred_class)

write.csv(confusion$table, here(paste0("Results/RF-confusiontable-", response, "-occurrence-2.csv")), row.names=F)

@@

# generate predictions for entire dataset
x <- datBB.c$PLT_CN
x <- cbind(x, predict(ranger_obj, datBB.c)$predictions[,2])
colnames(x) <- c("PLT_CN", "SBpred")
write.csv(x, here(paste0("Results/RF-pred-", response, "-occurrence-2.csv")), row.names=F)

x <- as.data.frame(ranger_obj$variable.importance)
write.csv(x, here(paste0("Results/RF-vimp-", response, "-occurrence-2.csv")), row.names=F)

# partial dependence
pd <- NULL
for(z in vars.final){
  pdz <- partial(ranger_obj, train=datz.train, pred.var=z, which.class=strsplit(response, split="agent")[[1]][2], prob=T)
  pdz$var <- z
  colnames(pdz) <- c("value","yhat","var")
  pd <- rbind(pd, pdz)
}
write.csv(pd,  here(paste0("Results/RF-partialdependence-", response, "-occurrence-2.csv")), row.names=F)
```

#### SB
```{r fit SB RF model, eval=F}
response <- "agentSB"
vars <- c("BA.PIEN", "QMD.PIEN", "propBA.PIEN")
datBB.c <- df[df$focalhost=="Any",]

splitit <- initial_split(datBB.c, prop=.7, strata=response)
datz.train <- training(splitit)
datz.test <- testing(splitit)
datz.train.cv <- vfold_cv(datz.train, strata=response, v=10)

# set formula
BBoccurrence_recipe <- recipe(formula(paste(response, "~", paste(vars, collapse = " + "))), datz.train)

# set sample weights
wts <- 1/(table(datBB.c[,response])/ max(table(datBB.c[,response])))
wts <- wts/sum(wts)

# set model
rf_model <- rand_forest() %>% set_args( trees=tune(), min_n=tune()) %>% set_engine("ranger", importance = "permutation", probability=T, seed=913, sample.fraction=wts) %>% set_mode("classification")

@@

# tune model
tgrid <- expand.grid(trees = c(100,500), min_n=c(10, 50, 100))
rf_tune_results <- rf_workflow %>% tune_grid(resamples = datz.train.cv,  grid = tgrid, metrics = metric_set(roc_auc, accuracy, sens, spec))

@@

# Use backward selection to identify important variables
#dfo.vsurf <- VSURF::VSURF(datz.train[,vars], factor(datz.train[,response]), RFimplem="ranger", parallel=T, ntree=param_best$trees, min.node.size=param_best$min_n)
#registerDoSEQ()
#vars.final <- unique(c(vars[dfo.vsurf$varselect.pred]))
vars.final <- vars
BBoccurrence_recipe_final <- recipe(formula(paste(response, "~", paste(vars.final, collapse = " + "))), data=datBB.c)

rf_model_final <- rand_forest() %>% set_args(trees=param_best$trees, min_n=param_best$min_n, splitrule="gini") %>% set_engine("ranger", importance = "permutation", probability=T, sample.fraction=wts) %>% set_mode("classification")

rf_workflow_final <- workflow() %>% add_recipe(BBoccurrence_recipe_final) %>% add_model(rf_model_final)
rf_workflow_final_fit<- rf_workflow_final %>% last_fit(splitit, metrics=metric_set(roc_auc, accuracy, sens, spec)) #emulates the process where, after determining the best model, the final fit on the entire training set is needed and is then evaluated on the test set. 
test_performance <- rf_workflow_final_fit %>% collect_metrics()
write.csv(test_performance, here(paste0("Results/RF-testperformance-", response, "-occurrence-2.csv")), row.names=F)

# generate a confusion matrix for test set
test_predictions <- rf_workflow_final_fit %>% collect_predictions()
confusion <- test_predictions %>% conf_mat(response,estimate = .pred_class)

write.csv(confusion$table, here(paste0("Results/RF-confusiontable-", response, "-occurrence-2.csv")), row.names=F)

@@

# generate predictions for entire dataset
x <- datBB.c$PLT_CN
x <- cbind(x, predict(ranger_obj, datBB.c)$predictions[,2])
colnames(x) <- c("PLT_CN", "SBpred")
write.csv(x, here(paste0("Results/RF-pred-", response, "-occurrence-2.csv")), row.names=F)

x <- as.data.frame(ranger_obj$variable.importance)
write.csv(x, here(paste0("Results/RF-vimp-", response, "-occurrence-2.csv")), row.names=F)

# partial dependence
pd <- NULL
for(z in vars.final){
  pdz <- partial(ranger_obj, train=datz.train, pred.var=z, which.class=strsplit(response, split="agent")[[1]][2], prob=T)
  pdz$var <- z
  colnames(pdz) <- c("value","yhat","var")
  pd <- rbind(pd, pdz)
}
write.csv(pd,  here(paste0("Results/RF-partialdependence-", response, "-occurrence-2.csv")), row.names=F)
```

#### WBBB
```{r fit WBBB RF model, eval=F}
response <- "agentWBBB"
vars <- c("BA.ABLA", "QMD.ABLA", "propBA.ABLA")
datBB.c <- df[df$focalhost=="Any",]

splitit <- initial_split(datBB.c, prop=.7, strata=response)
datz.train <- training(splitit)
datz.test <- testing(splitit)
datz.train.cv <- vfold_cv(datz.train, strata=response, v=10)

# set formula
BBoccurrence_recipe <- recipe(formula(paste(response, "~", paste(vars, collapse = " + "))), datz.train)

# set sample weights
wts <- 1/(table(datBB.c[,response])/ max(table(datBB.c[,response])))
wts <- wts/sum(wts)

# set model
rf_model <- rand_forest() %>% set_args( trees=tune(), min_n=tune()) %>% set_engine("ranger", importance = "permutation", probability=T, seed=913, sample.fraction=wts) %>% set_mode("classification")

@@

# tune model
tgrid <- expand.grid(trees = c(100,500), min_n=c(10, 50, 100))
rf_tune_results <- rf_workflow %>% tune_grid(resamples = datz.train.cv,  grid = tgrid, metrics = metric_set(roc_auc, accuracy, sens, spec))

@@

# Use backward selection to identify important variables
#dfo.vsurf <- VSURF::VSURF(datz.train[,vars], factor(datz.train[,response]), RFimplem="ranger", parallel=T, ntree=param_best$trees, min.node.size=param_best$min_n)
#registerDoSEQ()
#vars.final <- unique(c(vars[dfo.vsurf$varselect.pred]))
vars.final <- vars
BBoccurrence_recipe_final <- recipe(formula(paste(response, "~", paste(vars.final, collapse = " + "))), data=datBB.c)

rf_model_final <- rand_forest() %>% set_args(trees=param_best$trees, min_n=param_best$min_n, splitrule="gini") %>% set_engine("ranger", importance = "permutation", probability=T, sample.fraction=wts) %>% set_mode("classification")

rf_workflow_final <- workflow() %>% add_recipe(BBoccurrence_recipe_final) %>% add_model(rf_model_final)
rf_workflow_final_fit<- rf_workflow_final %>% last_fit(splitit, metrics=metric_set(roc_auc, accuracy, sens, spec)) #emulates the process where, after determining the best model, the final fit on the entire training set is needed and is then evaluated on the test set. 
test_performance <- rf_workflow_final_fit %>% collect_metrics()
write.csv(test_performance, here(paste0("Results/RF-testperformance-", response, "-occurrence-2.csv")), row.names=F)

# generate a confusion matrix for test set
test_predictions <- rf_workflow_final_fit %>% collect_predictions()
confusion <- test_predictions %>% conf_mat(response,estimate = .pred_class)

write.csv(confusion$table, here(paste0("Results/RF-confusiontable-", response, "-occurrence-2.csv")), row.names=F)

# generate predictions for entire dataset
x <- datBB.c$PLT_CN
x <- cbind(x, predict(ranger_obj, datBB.c)$predictions[,2])
colnames(x) <- c("PLT_CN", "SBpred")
write.csv(x, here(paste0("Results/RF-pred-", response, "-occurrence-2.csv")), row.names=F)

x <- as.data.frame(ranger_obj$variable.importance)
write.csv(x, here(paste0("Results/RF-vimp-", response, "-occurrence-2.csv")), row.names=F)

# partial dependence
pd <- NULL
for(z in vars.final){
  pdz <- partial(ranger_obj, train=datz.train, pred.var=z, which.class=strsplit(response, split="agent")[[1]][2], prob=T)
  pdz$var <- z
  colnames(pdz) <- c("value","yhat","var")
  pd <- rbind(pd, pdz)
}
write.csv(pd,  here(paste0("Results/RF-partialdependence-", response, "-occurrence-2.csv")), row.names=F)
```

```{r}
MPBpp <- read.csv(here("Results/RF-partialdependence-agentMPB-occurrence.csv"))
MPBpp$focalagent <- "MPB"
SBpp <- read.csv(here("Results/RF-partialdependence-agentSB-occurrence.csv"))
SBpp$focalagent <- "SB"
WBBBpp <- read.csv(here("Results/RF-partialdependence-agentWBBB-occurrence.csv"))
WBBBpp$focalagent <- "WBBB"

ppdat <- rbind(rbind(MPBpp, SBpp), WBBBpp)
ppdat$var[ppdat$var %in% paste0("BA.", c("PICO", "PIEN", "ABLA"))] <- "BA"
ppdat$var[ppdat$var %in% paste0("QMD.", c("PICO", "PIEN", "ABLA"))] <- "QMD"
ppdat$var[ppdat$var %in% paste0("propBA.", c("PICO", "PIEN", "ABLA"))] <- "propBA"

x <- df[!df$focalagent=="Any",]
x <- x[,c("QMD.host", "propTBA", "BA.host", "focalagent", "affected01")]
x$affected01 <- ifelse(x$affected01=="affected", 1, 0)

xm <- melt(x, id.vars=c("focalagent", "affected01"))
colnames(xm) <- c("focalagent", "yhat", "var", "value")
xm$var <- factor(xm$var, levels=c("QMD.host", "propTBA", "BA.host"), labels=c("QMD", "propBA", "BA"))
xm$focalagent <- as.character(xm$focalagent)
p <- ggplot(ppdat[ppdat$var=="BA",], aes(x=value, y=yhat, col=focalagent))+geom_smooth()+scale_color_manual(values=p1col)+geom_point(data=xm[xm$var=="BA",], size=0, )+ylim(0.1,0.6)
ggplot(xm[xm$var=="BA",],aes(x=value, y=yhat, col=focalagent)

ggplot(xm, aes(x=focalagent, y=value))
```

```{r}

mpb <- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="MPB","affected"], bbo.dat[bbo.dat$focalagent=="MPB","hostspred"]))
sb <- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="SB","affected"], bbo.dat[bbo.dat$focalagent=="SB","hostspred"]))
wbbb <- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="WBBB","affected"], bbo.dat[bbo.dat$focalagent=="WBBB","hostspred"]))
any<- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="Any","affected"], bbo.dat[bbo.dat$focalagent=="Any","hostspred"]))

>>>>

### Import libraries
ipak(c('tidyverse', 'tidymodels', 'reshape2', 'here', 'rgdal', 'sp', 'raster', 'stars', 'sf', 'prism', 'tigris', "tmap", 'fasterize', 'ggplot2','RColorBrewer','ggsci', 'patchwork', 'colorspace', 'grid', 'knitr',  'kableExtra', 'bookdown',  'pdp', 'ranger', 'VSURF', 'foreach'))

@@

### Under what stand conditions are bark beetle outbreaks likely?
```{r, eval=T}
datBB.c <- df[df$focalhost=="Any",]
splitit <- initial_split(datBB.c, prop=0.5)
datz.train <- training(splitit)
datz.test <- testing(splitit)
datz.train.cv <- vfold_cv(datz.train, v=10)
```

#### MPB
```{r fit MPB RF model, eval=F}
response <- "agentMPB"
vars <- c("BA.PICO", "QMD.PICO", "propBA.PICO")

form <- paste(response, "~", paste(vars, collapse = " + "))

wts <- 1/(table(datBB.c[,response])/ max(table(datBB.c[,response])))
wts <- wts/sum(wts)

# set formula
BBoccurrence_recipe <- recipe(formula(paste(response, "~", paste(vars, collapse = " + "))), datz.train) %>%step_corr(all_predictors())%>% prep()

# set model
rf_model <- rand_forest() %>% set_args(mtry = tune(), trees=tune(), min_n=tune(), sample.fraction=wts) %>% set_engine("ranger", importance = "permutation", probability=T) %>% set_mode("classification")

@@

# tune model
tgrid <- expand.grid(mtry = c(1,3), trees = c(100,500), min_n=c(50,100,500))
rf_tune_results <- rf_workflow %>% tune_grid(resamples = datz.train.cv,  grid = tgrid, metrics = metric_set(roc_auc, accuracy))

@@

# Use backward selection to identify important variables
dfo.vsurf <- VSURF::VSURF(datz.train[,vars], factor(datz.train[,response]), RFimplem="ranger", parallel=T, mtry=param_best$mtry, ntree=param_best$trees, min.node.size=param_best$min_n)
registerDoSEQ()
vars.final <- unique(c(vars[dfo.vsurf$varselect.pred]))

BBoccurrence_recipe_final <- recipe(formula(paste(response, "~", paste(vars.final, collapse = " + "))), data=datBB.c)

rf_model_final <- rand_forest() %>% set_args(mtry = param_best$mtry, trees=param_best$trees, min_n=param_best$min_n, splitrule="gini", probability=T) %>% set_engine("ranger",  sample.fraction=wts,  importance = "permutation", probability=T) %>% set_mode("classification")

rf_workflow_final <- workflow() %>% add_recipe(BBoccurrence_recipe_final) %>% add_model(rf_model_final)
rf_workflow_final_fit<- rf_workflow_final %>% last_fit(splitit) #emulates the process where, after determining the best model, the final fit on the entire training set is needed and is then evaluated on the test set. 
test_performance <- rf_workflow_final_fit %>% collect_metrics()
write.csv(test_performance, here("Results/RF-testperformance-MPB-occurrence.csv"), row.names=F)

# generate a confusion matrix for test set
test_predictions <- rf_workflow_final_fit %>% collect_predictions()
confusion <- test_predictions %>% conf_mat(response,estimate = .pred_class)
write.csv(confusion$table, here("Results/RF-confusiontable-MPB-occurrence.csv"), row.names=F)

@@

# generate predictions for entire dataset
x <- datBB.c$PLT_CN
x <- cbind(x, predict(ranger_obj, datBB.c)$predictions[,2])
colnames(x) <- c("PLT_CN", "MPBpred")
write.csv(x, here("Results/RF-pred-MPB-occurrence.csv"), row.names=F)

x <- as.data.frame(ranger_obj$variable.importance)
x$var <- row.names(x)
x$prop <- x[,1]/sum(x[,1])*100
x <- x[order(x$prop),]
x$var <- factor(x$var, levels=x$var, ordered=T)
write.csv(x, here("Results/RF-vimp-MPB-occurrence.csv"), row.names=F)

# partial dependence
pd <- NULL
for(z in vars.final){
  pdz <- partial(ranger_obj, train=datz.train, pred.var=z, which.class="MPB")
  pdz$var <- z
  colnames(pdz) <- c("value","yhat","var")
  pd <- rbind(pd, pdz)
}
write.csv(pd,  here("Results/RF-partialdependence-MPB-occurrence.csv"), row.names=F)

pdz <- partial(ranger_obj, train=datz.train, pred.var=c("QMD.PICO", "BA.PICO"), which.class="MPB")

#### SB
```{r fit SB RF model, eval=F}
response <- "agentSB"
vars <- c("BA.PIEN", "QMD.PIEN", "propBA.PIEN")

form <- paste(response, "~", paste(vars, collapse = " + "))

wts <- 1/(table(datBB.c[,response])/ max(table(datBB.c[,response])))
wts <- wts/sum(wts)

# set formula
BBoccurrence_recipe <- recipe(formula(paste(response, "~", paste(vars, collapse = " + "))), datz.train) %>%step_corr(all_predictors())%>% prep()

# set model
rf_model <- rand_forest() %>% set_args(mtry = tune(), trees=tune(), min_n=tune(), sample.fraction=wts) %>% set_engine("ranger", importance = "permutation", probability=T) %>% set_mode("classification")

@@

# tune model
tgrid <- expand.grid(mtry = c(1,3), trees = c(100,500), min_n=c(50,100,500))
rf_tune_results <- rf_workflow %>% tune_grid(resamples = datz.train.cv,  grid = tgrid, metrics = metric_set( accuracy))

@@

# Use backward selection to identify important variables
dfo.vsurf <- VSURF::VSURF(datz.train[,vars], factor(datz.train[,response]), RFimplem="ranger", parallel=T, mtry=param_best$mtry, ntree=param_best$trees, min.node.size=param_best$min_n)
registerDoSEQ()
vars.final <- unique(c(vars[dfo.vsurf$varselect.pred]))

BBoccurrence_recipe_final <- recipe(formula(paste(response, "~", paste(vars.final, collapse = " + "))), data=datBB.c)

rf_model_final <- rand_forest() %>% set_args(mtry = param_best$mtry, trees=param_best$trees, min_n=param_best$min_n, splitrule="gini", probability=T) %>% set_engine("ranger",  sample.fraction=wts,  importance = "permutation", probability=T) %>% set_mode("classification")

rf_workflow_final <- workflow() %>% add_recipe(BBoccurrence_recipe_final) %>% add_model(rf_model_final)
rf_workflow_final_fit<- rf_workflow_final %>% last_fit(splitit) #emulates the process where, after determining the best model, the final fit on the entire training set is needed and is then evaluated on the test set.

# generate a confusion matrix for test set
test_predictions <- rf_workflow_final_fit %>% collect_predictions()
confusion <- test_predictions %>% conf_mat(response,estimate = .pred_class)
write.csv(confusion$table, here("Results/RF-confusiontable-SB-occurrence.csv"), row.names=F)

@@

# generate predictions for entire dataset
x <- datBB.c$PLT_CN
x <- cbind(x, predict(ranger_obj, datBB.c)$predictions[,2])
colnames(x) <- c("PLT_CN", "SBpred")
write.csv(x, here("Results/RF-pred-SB-occurrence.csv"), row.names=F)

x <- as.data.frame(ranger_obj$variable.importance)
x$var <- row.names(x)
x$prop <- x[,1]/sum(x[,1])*100
x <- x[order(x$prop),]
x$var <- factor(x$var, levels=x$var, ordered=T)
write.csv(x, here("Results/RF-vimp-SB-occurrence.csv"), row.names=F)

# partial dependence
pd <- NULL
for(z in vars.final){
  pdz <- partial(ranger_obj, train=datz.train, pred.var=z, which.class="SB")
  pdz$var <- z
  colnames(pdz) <- c("value","yhat","var")
  pd <- rbind(pd, pdz)
}
write.csv(pd,  here("Results/RF-partialdependence-SB-occurrence.csv"), row.names=F)
```

#### WBBB
```{r fit WBBB RF model, eval=F}
response <- "agentWBBB"
vars <- c("BA.ABLA", "QMD.ABLA", "propBA.ABLA")

form <- paste(response, "~", paste(vars, collapse = " + "))

wts <- 1/(table(datBB.c[,response])/ max(table(datBB.c[,response])))
wts <- wts/sum(wts)

# set formula
BBoccurrence_recipe <- recipe(formula(paste(response, "~", paste(vars, collapse = " + "))), datz.train) %>%
step_corr(all_predictors())%>% prep()

# set model
rf_model <- rand_forest() %>% set_args(mtry = tune(), trees=tune(), min_n=tune(), sample.fraction=wts) %>% set_engine("ranger", importance = "permutation", probability=T) %>% set_mode("classification")

@@

# tune model
tgrid <- expand.grid(mtry = c(1,3), trees = c(100,500), min_n=c(30,100,500))
rf_tune_results <- rf_workflow %>% tune_grid(resamples = datz.train.cv,  grid = tgrid, metrics = metric_set(roc_auc, accuracy))

@@

# Use backward selection to identify important variables
dfo.vsurf <- VSURF::VSURF(datz.train[,vars], factor(datz.train[,response]), RFimplem="ranger", parallel=T, mtry=max(param_best$mtry,length(vars.final)), ntree=param_best$trees, min.node.size=param_best$min_n)
registerDoSEQ()
vars.final <- unique(c(vars[dfo.vsurf$varselect.pred]))

BBoccurrence_recipe_final <- recipe(formula(paste(response, "~", paste(vars.final, collapse = " + "))), data=datBB.c)

#BBoccurrence_recipe_final <-  recipe(formula("agentWBBB~1"), data=datBB.c)

rf_model_final <- rand_forest() %>% set_args(mtry = param_best$mtry, trees=param_best$trees, min_n=param_best$min_n, splitrule="gini", probability=T) %>% set_engine("ranger", sample.fraction=wts,  importance = "permutation", probability=T) %>% set_mode("classification")

rf_workflow_final <- workflow() %>% add_recipe(BBoccurrence_recipe_final) %>% add_model(rf_model_final)
rf_workflow_final_fit<- rf_workflow_final %>% last_fit(splitit)
#emulates the process where, after determining the best model, the #final fit on the entire training set is needed and is then evaluated #on the test set.

## generate a confusion matrix for test set
test_predictions <- rf_workflow_final_fit %>% collect_predictions()
confusion <- test_predictions %>% conf_mat(response,estimate = .pred_class)
write.csv(confusion$table, here("Results/RF-confusiontable-WBBB-occurrence.csv"), row.names=F)

## variable importance
final_model <- fit(rf_workflow_final, data=datz.train)
ranger_obj <- pull_workflow_fit(final_model)$fit

## generate predictions for entire dataset
x <- datBB.c$PLT_CN
x <- cbind(x, predict(ranger_obj, datBB.c)$predictions[,2])
colnames(x) <- c("PLT_CN", "WBBBpred")
write.csv(x, here("Results/RF-pred-WBBB-occurrence.csv"), row.names=F)

## variable importance
final_model <- fit(rf_workflow_final, data=datz.train)
ranger_obj <- pull_workflow_fit(final_model)$fit

x <- as.data.frame(ranger_obj$variable.importance)
x$var <- row.names(x)
x$prop <- x[,1]/sum(x[,1])*100
x <- x[order(x$prop),]
x$var <- factor(x$var, levels=x$var, ordered=T)
write.csv(x, here("Results/RF-vimp-WBBB-occurrence.csv"), row.names=F)

## partial dependence
pd <- NULL
for(z in vars.final){
  pdz <- partial(ranger_obj, train=datz.train, pred.var=z, which.class="WBBB")
  pdz$var <- z
  colnames(pdz) <- c("value","yhat","var")
  pd <- rbind(pd, pdz)
}
write.csv(pd,  here("Results/RF-partialdependence-WBBB-occurrence.csv"), row.names=F)
```

```{r}
wbbbp <- read.csv(here("Results/RF-partialdependence-WBBB-occurrence.csv"))

sbp <- read.csv(here("Results/RF-partialdependence-SB-occurrence.csv"))

mpbp <- read.csv(here("Results/RF-partialdependence-MPB-occurrence.csv"))

```{r, eval=F}
mpbo <- read.csv(here("Results/RF-pred-MPB-occurrence.csv"))
sbo <- read.csv(here("Results/RF-pred-SB-occurrence.csv"))
wbbbo <- read.csv(here("Results/RF-pred-WBBB-occurrence.csv"))

threshold <-0.5

bbo <- merge(mpbo, sbo, by="PLT_CN")
bbo <- merge(bbo, wbbbo, by="PLT_CN")

bbo$MPBpred01 <- ifelse(bbo$MPBpred>threshold,"MPB", "")
bbo$SBpred01  <- ifelse(bbo$SBpred>threshold,"SB", "")
bbo$WBBBpred01  <- ifelse(bbo$WBBBpred>threshold,"WBBB", "")

nrow(bbo[(bbo$MPBpred>threshold | bbo$SBpred> threshold | bbo$WBBBpred > threshold),])/nrow(bbo) # 97% of stands had conditions suitable for one agent

nrow(bbo[bbo$MPBpred>threshold ,])/nrow(bbo) # 55% of stands had conditions suitable for MPB

nrow(bbo[bbo$SBpred>threshold ,])/nrow(bbo) # 52% of stands had conditions suitable for SB

nrow(bbo[bbo$WBBBpred>threshold ,])/nrow(bbo) # 55% of stands had conditions suitable for WBBB

nrow(bbo[bbo$MPBpred>threshold & bbo$SBpred>threshold  ,])/nrow(bbo) # 55% of stands had conditions suitable for WBBB

nrow(bbo[bbo$SBpred>threshold & bbo$WBBBpred>threshold  ,])/nrow(bbo) # 55% of stands had conditions suitable for WBBB

nrow(bbo[bbo$SBpred>threshold & bbo$WBBBpred>threshold & bbo$MPBpred>threshold  ,])/nrow(bbo) # 55% of stands had conditions suitable for WBBB

bbo$agentspred <- do.call(paste, c(bbo[c("MPBpred01", "SBpred01", "WBBBpred01")],sep="-"))
bbo$agentspred <- factor(bbo$agentspred, level=c("--","MPB--","-SB-","--WBBB","MPB-SB-","MPB--WBBB","-SB-WBBB", "MPB-SB-WBBB"), labels=c("none", "MPB", "SB", "WBBB", "MPB & SB", "MPB & WBBB", "SB & WBBB", "MPB, SB, & WBBB"), ordered=T)

bbo$hostspred <- factor(bbo$agentspred , levels=c("none", "MPB", "SB", "WBBB", "MPB & SB", "MPB & WBBB", "SB & WBBB", "MPB, SB, & WBBB"),  labels=c("none", "PICO", "PIEN", "ABLA", "PICO\n& PIEN", "PICO\n& ABLA", "PIEN \n& ABLA", "PICO, PIEN,\n& ABLA"), ordered=T)

bbo.sub <- bbo[!bbo$agentspred == "none",   ]

bbo.dat <- merge(bbo.sub,df, by="PLT_CN", all=T)

```{r}
mpb <- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="MPB","affected"], bbo.dat[bbo.dat$focalagent=="MPB","hostspred"]))
sb <- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="SB","affected"], bbo.dat[bbo.dat$focalagent=="SB","hostspred"]))
wbbb <- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="WBBB","affected"], bbo.dat[bbo.dat$focalagent=="WBBB","hostspred"]))
any<- as.data.frame(table(bbo.dat[bbo.dat$focalagent=="Any","affected"], bbo.dat[bbo.dat$focalagent=="Any","hostspred"]))
